The human in the loop: volunteer computing as a socio-technical system

### 2016-02-29 Alife16/paper.tex
Starting with conclusion

In this paper two versions of a client-server architecture for volunteer and distributed
evolutionary algorithms that uses the browser and generated using {\sf
  NodIO}, based on the {\sf NodEO} evolutionary algorithm library have been
evaluated. Volunteers are {\em in the cloud}, as stated in the title,
since they are a {\em CPU as a service} for the persons running the
experiment. In fact, in this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative or in conjunction with other local or
cloud-based methods to add computing power in a seamless way through
the pool that NodIO creates.

In order to establish a baseline performance, the evolutionary
algorithm was run in a desktop client-program written in JavaScript
using NodEO to solve the 40-trap function. The first experiments with
{\sf NodIO} proved that, although obtaining better performance than the
baseline was possible, it did not happen, mainly because of the
difficulty in carrying over volunteers to other experiments when the
one they were participating on was finished, shown in the low
correlation between the number of IPs in successive experiments. This
also resulted in a low number of generations allotted by users.

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model depends on the implementation, with contributions following
a General Extreme Values distribution in the case of {\sf NodIO} and a
Weibull distribution for {\sf NodIO-W$^2$}. These distributions are
close enough to each other and, in the second case, reflect the fact
that all time spent in the page is actually devoted to computing,
which is why the time spent (represented by the number of
contributions to the pool) follows a model quite similar to that found
for games or other online activities. The reverse might be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they've done it
once, they might do it more times. In the spirit of Open Science, this
gamification might involve computing in real time data such as the one
presented in this paper and showing it in the same page.

>>>>

Our intention in this paper was to assess the capabilities of a
sociotechnical system formed by a client-server web-based framework
running a distributed evolutionary algorithms and the volunteers that
participate in the experiment. These volunteers are {\em in the
  cloud}, that is, available as {\em CPU as a service} for the persons
running the experiment. In this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative, or, if other computing resources are
available, in conjunction with other local or cloud-based methods to
add computing power in a seamless way through the pool that NodIO creates.

After running the experiment on the 40-trap problem whose running time
could be as low as a few seconds, we switched to another batch of
experiments where we used the 50-trap problem and also to a pool of
limited, and dwindling through the three experiments, size. Since our
initial results indicated that what happened on the screen, a flat
graph with no improvements or the experiment finished, influenced the
amount of time that the users devoted to the experiment, a longer one
could yield different results and, at the same time, result in a big
pool that might either crash the server or return useless individuals
to the volunteers. These new experiments have proved that using the
limited pool is beneficial to finding the solution, since less
evaluations are needed, but also that since the problem is more
difficult, the users stay for longer in the web page, making less
ephemeral the sociotechnical computing system created by the
simulation.

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model depends on the implementation, with contributions following
a Weibull distribution, which reflects the fact
that volunteer computing follows a model quite similar to that found
for games or other online activities. The reverse might be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they've done it
once, they might do it more times. In the spirit of Open Science, this
gamification might involve computing in real time data such as the one
presented in this paper and showing it in the same page or presenting
user results alongside others.
